---
title: "Calculating Growth Velocity"
author: "Hon Hwang"
date: "2017-08-31"
output:
  rmarkdown::html_vignette:
  number_sections: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

```{r init, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

library(hbgd)
```

The growth velocity function allows you to apply a function to the first
derivative of growth trajectories. For example, taking the mean of the
first derivative. Since the function depends on growth trajectories, they
need to be generated first; see
[documentation for HBGD package](https://hbgdki.github.io/hbgd/#fitting_all_trajectories).

```{r gen-trajectory, echo=FALSE, results='hide'}
smc <- get_smocc_data()[1:500,]
smc_haz_wandfit <- get_fit(smc, y_var = "haz", method = "wand")
smc_tr <- fit_all_trajectories(smc, smc_haz_wandfit)
```

# Calling the function

Suppose we assign the trajectories to a variable named `traj`. We call
the growth velocity method as follows:

```{r gen-velocity}
traj_velocity <- get_traj_velocity(all_traj = smc_tr)
```
The result of this call will be a
[tibble](https://cran.r-project.org/web/packages/tibble/index.html). Each row
of this `tibble` has the subject ID and growth velocity.

```{r inspect-velocity}
print(traj_velocity)
```

# Options and Defaults

* `all_traj` The returned value from calling `fit_all_trajectories`
  function. This is mandatory.
* `z_score = TRUE` will use the z-scores, e.g., `haz`. The default
   value is `TRUE`.
* `agedays_floor = 0.00` specifies the minimum `agedays` to use.
   The default value is zero.
   This parameter, combined with `agedays_limit`, extracts the required
   ranges of the fitted trajectory.
   Note that times depends heavily on the the number of grid points used to
   generate the trajectories, which is specified by the `xg` parameter for the
   functions `hbgd::fit_trajectory` and `hbgd::fit_all_trajectories`.
   This is because the function supplied by `FUN` will be applied to
   derivative data with x-axis grid value _less than_ `agedays_limit`; the
   x-axis grid values are set up by `xg` parameter of the
   functions `hbgd::fit_trajectory` and `hbgd::fit_all_trajectories`.   
* `agedays_limit = 365` specifies the maximum `agedays` to use.
   The default value is one year (365 days).
   This parameter, combined with `agedays_floor`, extracts the required
   ranges of the fitted trajectory.
   Note that times depends heavily on the the number of grid points used to
   generate the trajectories, which is specified by the `xg` parameter for the
   functions `hbgd::fit_trajectory` and `hbgd::fit_all_trajectories`.
   This is because the function supplied by `FUN` will be applied to
   derivative data with x-axis grid value _less than_ `agedays_limit`; the
   x-axis grid values are set up by `xg` parameter of the
   functions `hbgd::fit_trajectory` and `hbgd::fit_all_trajectories`.
* `FUN` argument takes a function specified by the user that will
  be apply on the growth trajectory data. This function should only
  receive one argument, a vector containing the first derivative
  of the growth trajectory. Since the first derivative can contain
  `NA` value (e.g., derivative for the first growth data), the supplied
  function should be able to handle `NA` values. The default for this
  argument is an internal function that simply calculates the mean.
